---
title: "Activity 13A - Hypothesis Tests for a Single Proportion"
author: "STS 2300 (Dr. VanKrevelen)"
date: 'Updated: 10/31/2023'
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Below are the packages you will need for this activity. If you use other packages, you can add them here.

```{r, message = FALSE}
library(ggplot2)
library(infer)
library(dplyr)
```

Let's suppose that we're interested in studying the concept of home field advantage in Major League Baseball during the 2020 season when fan attendance was restricted due to COVID-19. If there's no such thing as home field advantage, the home team will win just as often as the away team (half the time). If there *is* an advantage to the home team, they will win over half the time. To carry out this test, use a significance level of $\alpha = 0.10$.


<br>


1. Step 1 - Write out appropriate null and alternative hypotheses below. Define $p$ for this specific example.

**$H_0:$**

**$H_a:$**

**p =**

<br>


2. Step 2 - I gathered data from two random dates, which you can read into R below. Summarize this data in context of this problem.

```{r}
mlbgames <- data.frame(date = c(rep("8/21/20", 14), rep("9/7/20", 11)),
                       winner = c("H", "A", "A", "A", "H",
                                  "H", "H", "H", "H", "H",
                                  "H", "A", "A", "A",
                                  "H", "A", "A", "A", "H",
                                  "H", "H", "H", "H", "H",
                                  "H"))
```

$\hat{p} =$

<br>


3. Step 3 - Use the `infer` package to create a null distribution and shade in your p-value on the picture. Then calculate the p-value and write a sentence about what the p-value means for this specific example. Remove `eval = FALSE` when you're ready to run your code. (Note: You aren't giving a formal conclusion yet.)

```{r, eval = FALSE}
set.seed()
mlb_nulldist <- mlbgames %>%
  specify() %>%
  hypothesize() %H%
  generate() %>%
  calculate()
```

**p-value = **

**Sentence about p-value: **

<br>


4. Step 4 - Write a formal conclusion based on your p-value.

**Conclusion: **

<br>


5. Write a couple sentences about why you reached the conclusion you did. Do you think it's possible the conclusion is "wrong" (i.e. you rejected $H_0$ even though it's true or failed to reject it even though it's false)? What could we do to feel more confident in our answer?

**Answer:**

<br>


6. Try using the `prop.test()` function to carry out a theory-based hypothesis test. Do you get similar results?

```{r}

```

<br>


7. Want extra practice? Suppose the COVID-19 sniffing dog (from Activity 10) will only be effective if it can identify at least 75% of positive COVID-19 cases. The data below includes only the IDs the dog made for positive samples of COVID-19. Carry out a hypothesis test based on this example.

```{r}

covid_dog <- data.frame(ID = c(rep("positive", 157), rep("negative", 30)))
```

**Step 1 - Hypotheses**

**$H_0:$**

**$H_a:$**

**p =**

<br>



**Step 2 - Summarize data**

```{r}

```

$\hat{p} =$

<br>


**Step 3 - Compare data to $H_0$ with a null distribution and p-value**

```{r}

```

**p-value =**

<br>


**Step 4 - Formal conclusion**

**Conclusion:**