---
title: "Activity 14 - Hypothesis Tests for Means"
author: "STS 2300 (Dr. VanKrevelen)"
date: 'Updated: 11/4/2023'
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Below are the packages you will need for this activity. If you use other packages, you can add them here.

```{r, message = FALSE}
library(ggplot2)
library(infer)
library(dplyr)
```


# Example 1 - Prius Mileage

Suppose Toyota is claiming that their newest Prius Prime model gets over 100 miles per gallon on average. A car review website plans to test this claim. (Use $\alpha = 0.10$ for your significance level)

<br>

1. **Step 1** - Write out the appropriate hypotheses for this problem. Define any parameters below.

$H_0:$

$H_A:$

**Define parameter(s):**

<br>


2. **Step 2** - Gather and summarize data

The car review website has gathered data from 19 Prius Prime owners. The `average_mpg` variable is what each owner reported as their average miles per gallon over the past month. Below is the code to read in this data. 

Generate a graph to view the data and then calculate an appropriate statistic to summarize the data in relation to these hypotheses. Write a couple sentences about what your graph and statistic show you.

```{r}
prius <- read.csv("https://www.openintro.org/data/csv/prius_mpg.csv")
```

**Statistic:**

**Description:**

<br>


3. **Step 3** - Use the `infer` package to generate a null distribution. Visualize your p-value on the distribution and calculate the p-value. Write a sentence explaining what the p-value means in context. (Note: You are not writing a conclusion for your hypothesis test yet.)

```{r}
# Remember to set a seed
```

**p-value =**

**Sentence:**

<br>


4. **Step 4** - Write a conclusion for your hypothesis test Make sure that you address each of the important elements mentioned in class.

**Conclusion:**


<br>


5. What questions or concerns might we have before applying these conclusions too broadly? Consider what information you were given, how the data was collected, etc.

**Answer:**

<br>


---


# Example 2 - Fidget spinners

We previously looked at some data where people used fidget spinners and researchers measured their attention and memory. Suppose we want to compare the attention score between people who are near a fidget spinner (but not the one using it) and people with no spinner present. We suspect that those without a spinner present might have a higher memory score. (Use $\alpha = 0.05$ for your significance level)

<br>

1. **Step 1** - Write out the appropriate hypotheses for this problem. Define any parameters below.

$H_0:$

$H_A:$

**Define parameter(s)**

<br>


2. **Step 2** - Below is code to read in the data and to remove the third group (those actively using a spinner). Generate a graph to view the data and then calculate an appropriate statistic to summarize the data for these hypotheses. Write a couple sentences about what you notice.

```{r}
fidget <- read.csv("https://raw.githubusercontent.com/vank-stats/STS2300-Fall2023/main/fidget_spinners.csv") %>%
  filter(Condition != "Spinner")
```

**Statistic:**

**Description:**

<br>


3. **Step 3** - Use the `infer` package to generate a null distribution. Visualize your p-value on the distribution and calculate the p-value. Write a sentence explaining what the p-value means in context. (Note: You are not writing a conclusion for your hypothesis test yet.)

```{r}

```

**p-value =**

**Sentence:**

<br>


4. **Step 4** - Write a conclusion for your hypothesis test based on the p-value.

**Conclusion:**


<br>


5. Does this mean that fidget spinners aren't distracting for people nearby? If not, what does it mean?

**Answer:**


<br>


6. A 95% confidence interval for the difference in population means would be around (-0.2, 0.7). Discuss how this does or does not agree with your hypothesis test conclusion.

**Answer:**
