---
title: "Activity 15 - Project 2 Example"
author: "STS 2300 (Dr. VanKrevelen)"
date: 'Updated: 11/4/2023'
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Below are the packages you will need for this activity. If you use other packages, you can add them here.

```{r, message = FALSE}
library(ggplot2)
library(infer)
library(dplyr)
library(ipumsr)
```

For this activity, we are going to use Option 1 from the Project 2 data to see an example of a Chi-Squared test and an ANOVA test. This will give you an example of some things you might try / consider when working with this data.

**Before starting this activity, you will need to put the Option 1 data into the same folder as your .Rmd file**

<br>


# Data management

First, I don't want to use the full 100,000+ observations, so I'm going to subset my data to only include people who are between 30 and 65 years old in Alamance County. The age component will be important because I'm going to look at education levels and marriage status. If I included all ages, I would have very young people who haven't had time to do either of these things (or old people much more likely to be widowed).

```{r}
proj2 <- read_ipums_micro("usa_00012.xml")

ALAM <- filter(proj2, 
               COUNTYFIP == 1,
               AGE >= 30,
               AGE <= 65)
```



Next, I'm going to explore some of the variables I'm interested in. I want to study marital status (`MARST`) and education level obtained (`EDUC`). 

```{r}
table(ALAM$EDUC, ALAM$MARST)
```


Both variables have more categories than I reasonably want to work with. I'm going to combine several categories of each using `mutate()` and a function called `case_when()`. The `factor()` function is being used to put the categories in the proper order when I graph them.

```{r}
ALAM <- ALAM %>%
  mutate(education = case_when(EDUC < 6 ~ "Less than HS",
                               EDUC == 6 ~ "HS grad",
                               EDUC > 6 ~ "Attended college"),
         education = factor(education, levels = c("Less than HS",
                                                  "HS grad",
                                                  "Attended college")),
         marriage = case_when(MARST <= 2 ~ "Married",
                              MARST >2 & MARST < 6 ~ "Previously married",
                              MARST == 6 ~ "Never married"),
         marriage = factor(marriage, levels = c("Never married",
                                                "Married",
                                                "Previously married")))

table(ALAM$education, ALAM$marriage)
```


Suppose I'm also interested in how much people work per week. I'm going to subset my data to only include my variables of interest.

```{r}
ALAM <- ALAM %>%
  select(UHRSWORK, education, marriage)
```


# ANOVA Test

Let's do an ANOVA test to see if there are differences in population mean hours worked across education levels for 30 - 65 year olds in Alamance County.

1. What are the null and alternative hypotheses?

$H_0:$

$H_a:$

<br>


2. Summarize our data graphically and numerically.

```{r}

```

<br>


3. Generate a null distribution with a p-value shaded on the graph. (Remember to set a seed)

```{r}

```

<br>


4. Interpret your p-value in context of the problem. Then use your p-value to reach a formal conclusion.

**P-value interpretation:**

**Hypothesis test conclusion:**

<br>



# Chi-Squared Test

Let's do a Chi-Squared test to see if there is a relationship between education level and martial status among 30 - 65 year olds in Alamance County.

1. What are the null and alternative hypotheses?

$H_0:$

$H_a:$

<br>


2. Summarize our data graphically and numerically.

```{r}

```

<br>


3. Generate a null distribution with a p-value shaded on the graph. (Remember to set a seed)

```{r}

```

<br>


4. Interpret your p-value in context of the problem. Then use your p-value to reach a formal conclusion.

**P-value interpretation:**

**Hypothesis test conclusion:**

<br>



# Citing our data

Lastly, let's practice citing our data like you'll need to for the project. You can use the function `ipums_condition()` to get the citation infomation. Scroll down near the bottom of the output then copy and past the citation (statrting with "Steven Ruggles") and paste it into the document below.

**Include citation below**
