---
title: "Activity 03B - Joining Data Frames"
author: "STS 2300 (Dr. VanKrevelen)"
date: 'Updated: 8/29/2023'
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Load required packages here - change options to remove messages

library(dplyr)
library(Lahman) # You may need to install this one first
```

There is an online game called Immaculate Grid where people (i.e. baseball nerds) are asked to think of players who meet certain criteria. One example would be someone who played for the Minnesota Twins and recorded 200 hits in a season.

An R package called `Lahman` includes multiple data frames with different kinds of information that can help us practice our data wrangling along with the new skill of joining data frames.


<br>


1. Use a function like `str()` or `summary()` to explore the `Batting` data frame in the `Lahman` package. How many observations are there and how many variables?

```{r}

```

**Answer:**

<br>


2. There is a data set called `Batting` in the `Lahman` package. Write code to find all observations where someone played for the Minnesota Twins (variable `teamID` of `MIN`) and recorded at least 200 hits (variable `H`). Store the result in an object called `results`. How many rows are there? How many different players are there?

```{r}

```

**Answer:**

<br>


3. This output isn't very useful because most people won't know who `olivato01` is. However, we could find that out by using the `People` data frame. This data frame includes biographical information including a player's first and last name. We can use join functions to combine the information from our two data frames. When we use the function `left_join()` from the `dplyr` package, we take all of the observations in our first data frame and add information from the second data frame that matches a common key. Remove `eval = FALSE` and observe the results of the new data frame.

```{r, eval = FALSE}
left_join(results, People, by = join_by(playerID))
```

<br>


4. In addition to choosing which variables we keep from a data frame, the `select()` function also lets us choose an order. Take the code from the previous question and use `select()` so that we only see `nameFirst`, `nameLast`, `yearID`, `H`. Now we have something that might be useful to a person playing this game!

```{r}

```

<br>


5. Another function you might want to use is called `inner_join()`. This function will return only rows that have matching values in both data frames. Sometimes Immaculate Grid will ask people to name a player who played for two different teams. Suppose we want to determine hitters (`Batting`) data frame who played for the the Boston Red Sox (teamID of BOS) and the New York Yankees (teamID of NYA). Start by creating a data frame called `BOS` that only includes Red Sox players. Then summarize that data frame by player with a variable called `Seasons` that counts how many times each player shows up in the data.Then count the number of times a player shows up using `summarize()` and `.by = playerID`. Next do the same think for the Yankees and create `NYA`.

```{r}

```

<br>


6. Now do the same thing for the Yankees to create an object called `NYA`.

```{r}

```

<br>


7. Once you have both data frames, use `inner_join()` to join them by `playerID`. Your code will look very similar to when we used `left_join()`. What does your resulting data frame look like?

```{r}

```

<br>


8. At this point, we likely want to rename our Seasons.x and Seasons.y columns and add names  from the `People` data frame again. See if you can produce a data frame that lists only:

  - firstName
  - lastName
  - BOS_Seasons
  - NYA_Seasons
  
```{r}

```

**How many batters played for both teams?**:


<br>


**Note:** Want to learn more about joining data frames together?

  - [Section 3.7 of ModernDive](https://moderndive.com/3-wrangling.html#joins) includes examples using airline data
  - [Chapter 20 of R for Data Science](https://r4ds.hadley.nz/joins) uses similar examples
  - [Chapter 5 of Modern Data Science with R](https://mdsr-book.github.io/mdsr3e/05-joins.html) uses an airline example and a baseball example