---
title: "Activity 13A - Hypothesis Tests for a Single Proportion"
author: "STS 2300 (Dr. VanKrevelen)"
date: 'Updated: 10/31/2023'
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Below are the packages you will need for this activity. If you use other packages, you can add them here.

```{r, message = FALSE}
library(ggplot2)
library(infer)
library(dplyr)
```

Let's suppose that we're interested in studying the concept of home field advantage in Major League Baseball during the 2020 season when fan attendance was restricted due to COVID-19. If there's no such thing as home field advantage, the home team will win just as often as the away team (half the time). If there *is* an advantage to the home team, they will win over half the time. To carry out this test, use a significance level of $\alpha = 0.10$.


<br>


1. Step 1 - Write out appropriate null and alternative hypotheses below. Define $p$ for this specific example.

**$H_0: p = 0.5$**

**$H_A: p > 0.5$**

**p = the population proportion of all games that are won by the home team**

<br>


2. Step 2 - I gathered data from two random dates, which you can read into R below. Summarize this data in context of this problem.

```{r}
mlbgames <- data.frame(date = c(rep("8/21/20", 14), rep("9/7/20", 11)),
                       winner = c("H", "A", "A", "A", "H",
                                  "H", "H", "H", "H", "H",
                                  "H", "A", "A", "A",
                                  "H", "A", "A", "A", "H",
                                  "H", "H", "H", "H", "H",
                                  "H"))

mlb_phat <- mlbgames %>%
  specify(winner ~ NULL, success = "H") %>%
  calculate(stat = "prop")

mlb_phat
```

$\hat{p} = 0.64$

<br>


3. Step 3 - Use the `infer` package to create a null distribution and shade in your p-value on the picture. Then calculate the p-value and write a sentence about what the p-value means for this specific example. Remove `eval = FALSE` when you're ready to run your code. (Note: You aren't giving a formal conclusion yet.)

```{r, eval = TRUE}
set.seed(1031)

mlb_null <- mlbgames %>%
  specify(winner ~ NULL, success = "H") %>%
  hypothesize(null = "point", p = 0.5) %>%
  generate(reps = 1000, type = "draw") %>%
  calculate(stat = "prop")

visualize(mlb_null) +
  shade_pvalue(obs_stat = mlb_phat, 
               direction = "right")

get_p_value(mlb_null, 
            obs_stat = mlb_phat, 
            direction = "right")
```

**p-value =** 0.115

**Sentence about p-value:** In a world with no home field advantage, the home team would win 64% (our p-hat) or more (because of Ha) games in a sample of 25 games around 11.5% of the time.

<br>


4. Step 4 - Write a formal conclusion based on your p-value.

**Conclusion:** My p-value of 0.115 is greater than the significance level of 0.10, so my data is compatible with $H_0$ and I will fail to reject the null hypothesis. There is insufficient evidence to conclude home teams win more than half of the time in Major League Baseball games (in 2020).

<br>


5. Write a couple sentences about why you reached the conclusion you did. Do you think it's possible the conclusion is "wrong" (i.e. you rejected $H_0$ even though it's true or failed to reject it even though it's false)? What could we do to feel more confident in our answer?

**Answer:** It's entirely possible that there actually is a home field advantage. Remember that our conclusion says we "fail to reject" the hypothesis of no home field advantage, but that doesn't mean we are convinced it is true. We only looked at 25 games, and that data by itself was compatible with the null hypothesis. If we wanted to do a better investigation of this, it might make sense to take a bigger sample (say 100 games).

<br>


6. Try using the `prop.test()` function to carry out a theory-based hypothesis test. Do you get similar results?

```{r}
prop.test(x = 16, 
          n = 25, 
          p = 0.5, 
          alternative = "greater")
```

<br>


7. Want extra practice? Suppose the COVID-19 sniffing dog (from Activity 10) will only be effective if it can identify at least 75% of positive COVID-19 cases. The data below includes only the IDs the dog made for positive samples of COVID-19. Carry out a hypothesis test based on this example.

```{r}
covid_dog <- data.frame(ID = c(rep("positive", 157), rep("negative", 30)))
```

**Step 1 - Hypotheses**

**$H_0:$** p = 0.75

**$H_a:$** p > 0.75

**p =** the population proportion of times the dog can correctly identify positive COVID-19 cases

<br>



**Step 2 - Summarize data**

```{r}
covid_dog %>%
  specify(ID ~ NULL, success = "positive") %>%
  calculate(stat = "prop")
```

$\hat{p} = 0.840$ 

<br>


**Step 3 - Compare data to $H_0$ with a null distribution and p-value**

```{r}
set.seed(111023)

dog_null <- covid_dog %>%
  specify(ID ~ NULL, success = "positive") %>%
  hypothesize(null = "point", p = 0.75) %>%
  generate(reps = 1000, type = "draw") %>%
  calculate(stat = "prop")

visualize(dog_null) +
  shade_pvalue(obs_stat = 0.840,
               direction = "right")

get_pvalue(dog_null,
           obs_stat = 0.840,
           direction = "right")
```

**p-value =** 0.001

<br>


**Step 4 - Formal conclusion**

**Conclusion:** Because the p-value of 0.001 is less than the significance level of 0.05, we reject the null hypothesis. There is very strong evidence to conclude the dog can correctly identify more than 75% of positive COVID-19 cases (in a controlled setting like the one in which the data was collected)