---
title: "Activity 12 - The Taylor Swift Effect?"
author: "STS 2300 (Dr. VanKrevelen)"
date: 'Updated: 10/26/2023'
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Below are the packages you will need for this activity. If you use other packages, you can add them here.

```{r, message = FALSE}
library(ggplot2)
library(infer)
```

According to [People.com](https://people.com/travis-kelce-taylor-swift-improved-performance-nfl-8379432), Travis Kelce has been performing better when Taylor Swift attends his football games. From the article, Travis Kelce's brother says to him:

> “When T Swift is at the game, you are averaging 99 yards and when you are left with only your friends and remaining family there, you are at 46.5 yards a game.”

We might wonder if Taylor Swift's presence makes Travis Kelce perform better or if this is just random variation that people are talking about.

Below is the data we will use ([Source](https://docs.google.com/spreadsheets/d/1Vmhu2BSiKOMQjXTdZiq8Gq51oVu-FudqPe6Sqpv7GJs/edit#gid=0)). It includes whether or not Taylor Swift attended the game (`taylor`) and how many receiving yards Kelce had (`yards`).

```{r}
# Data updated through 10/22

kelce <- data.frame(taylor = c("yes", "yes", "no", "yes", "yes", "no"),
                    yards = c(179, 124, 67, 60, 69, 26))
```

1. Step 1 - What are the null and alternative hypotheses for the claims being made here?

$H_0$: Taylor Swift's presence has no impact on how many receiving yards Travis Kelce gets

$H_a$: Taylor Swift's presence increases the receiving yards Travic Kelce gets

<br>


2. Step 2 - Calculate an appropriate statistic for this data. In addition, make a graph showing Kelce's receiving yards with and without Taylor Swift in attendance. What do you notice? (Note: Want to make your graph extra fancy? Consider using the [`tayloRswift` package.](https://cran.r-project.org/web/packages/tayloRswift/readme/README.html))

```{r}
obs_diff <- kelce %>%
  specify(yards ~ taylor) %>%
  calculate(stat = "diff in means", order = c("yes", "no"))
obs_diff

ggplot(kelce, aes(x = yards, y = taylor)) +
  geom_boxplot(fill = "red") +
  geom_point(color = "yellow", size = 3)
```

**Statistic:** $\bar{x}_1 - \bar{x}_2$ = 61.5

**Graph comments:** While there are only two games that Taylor Swift didn't attend, they are two of the three games Travis Kelce recorded his lowest number of receiving yards. His two best games were with Taylor Swift in attendance.

<br>


3. Step 3 - Let's generate a null distribution to explore what Travis Kelce's receiving yards would look like in a world where Taylor Swift had no impact on his performance. Graph your null distribution (with your p-value shaded) and comment on what it shows. Calculate the p-value as well.

```{r}
set.seed(10107)
swift_null <- kelce %>%
  specify(yards ~ taylor) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("yes", "no"))

visualize(swift_null) +
  shade_pvalue(obs_stat = obs_diff,
               direction = "right")

swift_null %>%
  get_pvalue(obs_stat = obs_diff,
             direction = "right")
```

**Null distribution comments:** The observed difference is near the far right side. An area of 0.135 is shaded in, which tells us that while Travis Kelce has performed better with Taylor Swift at his games so far, there is still a reasonable chance of this occurring due to random chance. There are some gaps in the null distribution because with only six observations there are only so many possible differences in means.

**p-value =** 0.135

<br>


4. Step 3 - Briefly describe how the null distribution is created using permutation. (Hint: Imagine that you have slips of paper with the number of yards for each game in front of you. How would you use those to create the null distribution?)

**Answer:** When we use `type = "permute"`, we are essentially shuffling up the slips of paper with his yards in each game and then reassigning them to the `taylor` variable. Each time we do this, we calculate a difference in means to see what our data would look like in a world where his yards are not dependent at all on whether Taylor Swift attends the game.

<br>


5. Step 4 - Use your p-value to write a conclusion for your hypothesis test in context of this example.

**Conclusion:** With a p-value of 0.135 that is greater than 0.05, we fail to reject the null hypothesis. There is insufficient evidence (after only six games) to conclude that Taylor Swift's presence increases the receiving yards Travic Kelce gets. 

<br>


6. Do you think it would make sense to use a theory-based hypothesis test to answer this question? Why or why not? (Hint: It would be the same test / conditions as we discussed with confidence intervals for a difference in means)

**Answer:** For a theory-based test for a difference in means, we would need each sample to come from a population with a normal distribution or for the sample sizes to be "large enough". Both groups have a VERY small sample size. We don't have much information about whether the population is normal, so we should be very wary of using theory-based methods for this problem.

